OECDactuals

### 2024-05-08 manuscript/compare_oecd_imf.Rmd
comparison of scores - oecd and imf actuals

---
title: "OECD Actuals and BVAR Specifications"
author: "Rike"
date: "2024-04-18"
output: pdf_document
---

@@

```{r cars, echo =FALSE, message=FALSE}
library(here)
library(data.table)
library(ggplot2)
library(MetBrewer)
library(patchwork)

@@

### OECD actuals vs. IMF actuals

@@

```{r, echo = FALSE}
allscores <- imfimf |>
  .d(, relwis := IMF - bvar_qu) |>
  dcast(country + target + horizon ~ truth, value.var = "relwis") |>
  .d(,horizon := factor(horizon, levels = c(0, 0.5, 1, 1.5), labels = 
                         c("Fall, same year", "Spring, same year",
                           "Fall, next year", "Spring, next year")))

ggplot(aes(x = imf, y = oecd, color = as.factor(horizon)), data = allscores) +
  geom_point() +
  geom_hline(aes(yintercept = 0)) +
  geom_vline(aes(xintercept = 0)) +
  scale_color_met_d("VanGogh2") +
  theme_uqimf() +
  scale_x_continuous(limits =  c(-0.55, 0.55)) +
  scale_y_continuous(limits =  c(-0.55, 0.55)) +
  ylab("OECD truth") +
  xlab("IMF truth")

## GDP separate

```{r, echo = FALSE, fig.dim=c(7,4)}
allscores <- imfimf |>
  .d(, relwis := IMF/bvar_qu) |>
  dcast(country + target + horizon ~ truth, value.var = "relwis") |>
  .d(,horizon := factor(horizon, levels = c(0, 0.5, 1, 1.5), labels = 
                         c("Fall, same year", "Spring, same year",
                           "Fall, next year", "Spring, next year")))

ggplot(aes(x = imf, y = oecd, color = as.factor(horizon)), data = allscores |> .d(target == "ngdp_rpch")) +
  geom_point() +
  geom_hline(aes(yintercept = 1)) +
  geom_vline(aes(xintercept = 1)) +
  scale_color_met_d("VanGogh2") +
  theme_uqimf() +
  scale_x_continuous(limits =  c(0, 5)) +
  ylab("OECD truth") +
  xlab("IMF truth")

## GDP separate

```{r, echo = FALSE, fig.dim=c(7,4)}
allscores <- imfimf |>
  .d(, relwis := IMF/bvar_qu) |>
  dcast(country + target + horizon ~ truth, value.var = "relwis") |>
  .d(,horizon := factor(horizon, levels = c(0, 0.5, 1, 1.5), labels = 
                         c("Fall, same year", "Spring, same year",
                           "Fall, next year", "Spring, next year")))

ggplot(aes(x = imf, y = oecd, color = as.factor(horizon)), data = allscores |> .d(target == "pcpi_pch")) +
  geom_point() +
  geom_hline(aes(yintercept = 1)) +
  geom_vline(aes(xintercept = 1)) +
  scale_color_met_d("VanGogh2") +
  theme_uqimf() +
  scale_x_continuous(limits =  c(0, 5)) +
  ylab("OECD truth") +
  xlab("IMF truth")

@@

```{r overall_assessment_ew, fig.dim=c(12,12), echo = FALSE, warning = FALSE, fig.height=5, fig.width=9}

@@

  dir_wis_plot <- wis_plot_new(decomp_data |> .d(error_method  == "directional"), trgt, xmax = xmax, plot_name = "WIS - Directional Errors", metcolor = "VanGogh2")
  abs_wis_plot <- wis_plot_new(decomp_data |> .d(error_method  == "absolute"), trgt, xmax = xmax, plot_name = "WIS - Absolute Errors", metcolor = "VanGogh2")

@@

  ovr_plot <- 
    (abs_wis_plot) +
    plot_layout(guides = "collect", 
                heights = c(0.5, 0.5)) &
      plot_annotation(tag_levels = 'I')  &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin=margin())

@@

## GDP

@@

## Histogram 
Values < 1 mean that the CISS-BVAR is better

```{r, echo=FALSE}
bvar_twospec <- fread(here("scores", "bvar_ci_scores.csv")) |>
  .d(model %in% c("bvar_qu", "bvar_ciss")) |>
  dcast(target + horizon  + country ~ model, value.var = "interval_score") |>
  .d(!is.na(bvar_ciss)) |> # no data for CAN, JPN 
  .d(, relwis := bvar_ciss/bvar_qu)

ggplot(aes(x = relwis), data = bvar_twospec) +
  geom_histogram(binwidth = 0.13345, color = "steelblue4", fill = "steelblue3") +
  theme_uqimf() +
  xlab("Relative WIS (CISS.BVAR / Reg.BVAR)")
```

>>>>

---
title: "OECDactuals"
author: "Rike"
date: "2024-04-17"
output: html_document
---

@@

```{r cars, echo =FALSE}
library(here)
library(data.table)
library(ggplot2)
library(MetBrewer)
library(patchwork)

@@

##

@@

```{r, echo = FALSE}
allscores <- imfimf |>
  .d(, relwis := IMF/bvar_qu) |>
  dcast(country + target + horizon ~ truth, value.var = "relwis") |>
  .d(,horizon := factor(horizon, levels = c(0, 0.5, 1, 1.5), labels = 
                         c("Fall, same year", "Spring, same year",
                           "Fall, next year", "Spring, next year")))

ggplot(aes(x = imf, y = oecd, color = as.factor(horizon)), data = allscores) +
  geom_point() +
  geom_hline(aes(yintercept = 1)) +
  geom_vline(aes(xintercept = 1)) +
  scale_color_met_d("VanGogh2") +
  theme_uqimf() +
  scale_x_continuous(limits =  c(0, 5)) +
  ylab("OECD truth") +
  xlab("IMF truth")

@@

```{r overall_assessment_ew, fig.dim=c(12,12), echo = FALSE, warning = FALSE}

@@

  dir_wis_plot <- wis_plot_new(decomp_data |> .d(error_method  == "directional"), trgt, xmax = xmax, plot_name = "WIS - Directional Errors", metcolor = "VanGogh2")
  abs_wis_plot <- wis_plot_new(decomp_data |> .d(error_method  == "absolute"), trgt, xmax = xmax, plot_name = "WIS - Absolute Errors")

@@

  ovr_plot <- 
    (abs_wis_plot) +
    plot_layout(guides = "collect", 
                heights = c(0.2, 0.05, 0.2, 0.05, 0.2)) &
      plot_annotation(tag_levels = 'I')  &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin=margin())

@@

```
## GDP

@@

## Histogram

```{r}
bvar_twospec <- fread(here("scores", "bvar_ci_scores.csv")) |>
  .d(model %in% c("bvar_qu", "bvar_ciss")) |>
  dcast(target + horizon  + country ~ model, value.var = "interval_score") |>
  .d(!is.na(bvar_ciss)) |> # no data for CAN, JPN 
  .d(, relwis := bvar_ciss/bvar_qu)

ggplot(aes(x = relwis), data = bvar_twospec) +
  geom_histogram(binwidth = 0.13345, color = "steelblue4", fill = "steelblue3") +
  theme_uqimf() +
  xlab("Relative WIS (Reg.BVAR/CISS.BVAR)")
```
